---
title: "Security Policy Forecasts"
author: "SPFT"
date: "24 Februar 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Clear Global environment
rm(list=ls())

## Setting Working directory
try(setwd("D:/Eigene Datein/Dokumente/Uni/Hertie/Materials/Master thesis/SecurityPolicyForecastingTournament"), silent = TRUE)

source("main.R")
```

## Security Policy Forecasting Tournament

- 53 Hertie students participated
(48 on time, 5 late)

- 24 forecasting questions 

- time horizon: February 12 to April 24, 2017


## Different representations forecasts
```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(1)

response.hertie2(1)
```

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.all(1)

response.all2(1)
```

## Question 1-4

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(1)

response.hertie(2)
```

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(3)

response.hertie(4)
```

## Question 5-8

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(5)

response.hertie(6)
```

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(7)

response.hertie(8)
```

## Question 9-12

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(9)

response.hertie(10)
```

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(11)

response.hertie(12)
```

## Question 13-16

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(13)

response.hertie(14)
```

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(15)

response.hertie(16)
```

## Question 17-20

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(17)

response.hertie(18)
```

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(19)

response.hertie(20)
```

## Question 21-24

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(21)

response.hertie(22)
```

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.hertie(23)

response.hertie(24)
```

## Control Variables: H1

Hypothesis 1a: More intelligent individuals are more accurate forecasters 

Hypothesis 1b: Higher moral competency does not improve the accuracy of forecasters

Idea: Moral thinking collides with analytical thinking (Tetlock, Gardner 2015, p. 229)

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
bnt.plot
mct.plot
```

## Control Variables: H2

Hypothesis 2: The more time forecasters use for forecasting the more accurate their forecast become

```{r, echo = FALSE, fig.width=7, fig.height=4, out.extra='style="float:left"'}
time.plot
```

## Control Variables: H3

Hypothesis 3:	Mild interventions encouraging the use of an analytical approach increase the accuracy of forecasting

All participants are assigned randomly in about equal shares to a treatment or control group. The treatment group is provided with a short (150 words) information about the way many successful forecasters would approach the forecasting question. 

```{r, echo = FALSE, fig.width=4, fig.height=3, out.extra='style="float:center"'}
group.plot
```

## Backup Hypotheses HA

Hypothesis: Forecasters using of a larger variety of information sources are more accurate.

```{r, echo = FALSE, fig.width=4, fig.height=4, out.extra='style="float:center"'}
source.plot
source.var.plot
```

## Descriptive Statistics

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
emp.plot
exp.plot
pop.plot
edu.plot
```