---
title: "Forecasting in IR"
subtitle: "Security Policy Forecasting Tournament"
author: "Alexander Sacharow"
date: "02 May 2017"
output: 
  ioslides_presentation:
    css: general/tables.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Clear Global environment
rm(list=ls())

## Setting Working directory
try(setwd("D:/Eigene Datein/Dokumente/Uni/Hertie/Materials/Master thesis/SecurityPolicyForecastingTournament"), silent = TRUE)

library(knitr)

source("main.R")
```

## Why Forecasting? {.text40}

<div class="centered">
What do we want to achieve with forecasting?
</div>

## Why Forecasting?


<div class="columns-2">

<div style="color:#C02F39;">**Authors: Scientists, Analysts**</div>

- to nail it

    - value in itself?

- test theories (Friedman, 1966; Ward, 2016)

- Central function of science next to explanation

    - how much can we know?

<div style="color:#C02F39;">**Users: Policy makers etc.**</div>

- basis of decision making

    - influence future

    - show alternatives

- Early-Warning

    - Make preparation possible
</div>

## Is Forecasting Possible?

<div class="centered">

<p style="font-size: 40px">
Is successful forecasting in international politics possible at all?
</p>

</div>


## Is Forecasting Possible?

<div class="columns-2">

<div style="color:#C02F39;">**The Skeptics**</div>

- Politics are like *Clouds* (Almond et al., 1977)

- Non-Linearity
   
    - limit to conclusions from the past
   
- *Black swan* (Taleb, 2007)
   
    - High impact events are often unexpected
   
    - e.g. discovery of America, Arab Spring

<div style="color:#C02F39;">**The Optimists**</div>

- *The poets*
   
    - Yes, but without probabilities (Scenarios, Mega trends)
  
- *The mathematicians*

    - Yes, with sufficient data (Chourci, 1974; Ward, 2016)
  
- With crowds
  
    - Groups, competitions, markets

</div>

## Forecasting methods - overview

1. **Strategic Foresight**

2. **Statistical models**

3. **Forecasting competitions**

4. Intuitive forecasts

5. Expert forecasts

6. Prediction markets

7. Intelligence Estimates

## Strategic Foresight

Example: [European Union in the World 2025](http://www.dahrendorf-forum.eu/wp-content/uploads/2016/05/Dahrendorf_Analysis_EU_2025_15.pdf) by Dahrendorf Forum

- Essentially: Scenario analysis

- originated  and widely used in the corporate sector

    - Shell Foresight 1960s (Pierre Wack)

<div style="color:#C02F39;">**Advantage:**</div>

- No content restrictions

- consistency of scenarios as key quality criteria

<div style="color:#C02F39;">**Disadvantage:**</div> 

- No likelihood judgements (policy maker ask for it)

- how can scenarios be verify?

## Statistical Models

Example: *World-Wide Integrated Crisis Early Warning System (W-ICEWS)* [(Preview)](http://nbviewer.jupyter.org/gist/dmasad/f79ce5abfd4fb61d253b)

<div style="color:#C02F39;">**Advantage:**</div>

- can be verified

- less decision biases

<div style="color:#C02F39;">**Disadvantage:**</div>

- requires measureable events

    - e.g. number of injured / death
    
- not applicable to many relevant questions

    - e.g. International Court of Justice decision forecast
    
- data often not existing / too expensive

## Forecasting Competitions

Example: [The Good Judgement Project](www.gjopen.com)

- for geopolitics relatively new (first in 2011)

**Security Policy Forecasting Tournament**

- 230 participants in total

    - 53 Hertie students participated (48 on time, 5 late)

- short-term: 2.5 months

## Performance

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.uni(9) + geom_vline(xintercept=0,  colour="blue", linetype="dashed", size=1.5)

response.uni(16) + geom_vline(xintercept=1,  colour="blue", linetype="dashed", size=1.5)
```

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.uni(13) + geom_vline(xintercept=0,  colour="blue", linetype="dashed", size=1.5)

response.uni(6) + geom_vline(xintercept=0,  colour="blue", linetype="dashed", size=1.5)
```

## The research in die background

**Hypothesis 2: The marginal added value of time spend on forecasting is positive and decreases over time**

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
cor.brier.time.log.plot
```

## Hertie Group compared to others

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
ggplot(SB, aes(x = brier.avg, fill = part.group)) +
     geom_histogram(binwidth=.1, position="dodge") + # bar type
     #geom_density(alpha=.3) +
     theme_bw() +
     theme(axis.title = element_text(size=18, colour = "#696969", family = "serif"), # Labels axis font size
           axis.text = element_text(size=14, colour = "#696969"),
           axis.line = element_line(colour = "#696969"), 
           axis.ticks = element_line(colour = "#696969")) +  
     labs( # title = "Brier score distribution",
          x = "Brier score",
          y = "Frequency") # labels))
```


<!--
## Different representations forecasts
```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.uni(1)

response.uni2(1)
```

```{r, echo = FALSE, fig.width=4, fig.height=2.6, out.extra='style="float:left"'}
response.all(1)

response.all2(1)
```


-->
## Scoreboard

```{r, echo = FALSE, message=FALSE, warning=FALSE}

SB %>% merge(read.csv("personalData.csv")[, c("ResponseId", "id.uni")], 
                      by = "ResponseId") %>% filter(part.group == "uni") %>% arrange(brier.avg) %>% 
     select(HertieID = id.uni, brier.avg) %>%  head(10) %>% mutate(rank = row_number()) %>% kable(format = "html")
```

